---
title: "Tidycensus"
author: "Karen Torres"
date: "2022-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidycensus)
options(tigris_use_cache = TRUE)

```

```{r}
total_population_10 <- get_decennial(
  geography = "state", 
  variables = "P001001",
  year = 2010)
total_population_10
```
```{r}
dc_income <- get_acs(
  geography = "tract", 
  variables = "B19013_001",
  state = "DC", 
  year = 2020,
  geometry = TRUE)
dc_income

```

```{r}
plot(dc_income["estimate"])

```
```{r}
boston <-get_acs(geography= "tract",
  variables = "B19013_001",
  state = "MA", 
  year = 2020,
  county = 025,
  geometry = TRUE)
boston
```

```{r}
plot(boston["estimate"])
```

```{r}
library(tidycensus)
library (tidyverse)
hennepin_race <- get_decennial(
  geography = "tract",
  state = "MN",
  county = "Hennepin",
  variables = c(
    Hispanic = "P2_002N",
    White = "P2_005N",
    Black = "P2_006N",
    Native = "P2_007N",
    Asian = "P2_008N"
  ),
  summary_var = "P2_001N",
  year = 2020,
  geometry = TRUE
) |>
  mutate(percent = 100 * (value / summary_value))
hennepin_race
```

```{r}
library(tmap)
hennepin_black <- filter(hennepin_race, 
                         variable == "Black")

tm_shape(hennepin_black) + 
  tm_polygons()


```{r}
tm_shape(hennepin_black) + 
  tm_polygons(col = "percent")
```

```{r}
tm_shape(hennepin_black) + 
  tm_polygons(col = "percent",
          style = "quantile",
          n = 5,
          palette = "Purples",
          title = "2020 US Census") + 
  tm_layout(title = "Percent Black\nby Census tract",
            frame = FALSE,
            legend.outside = TRUE)
```
```{r}
tm_shape(hennepin_black) + 
  tm_polygons(col = "percent",
          style = "jenks",
          n = 5,
          palette = "Purples",
          title = "2020 US Census",
          legend.hist = TRUE) + 
  tm_layout(title = "Percent Black\nby Census tract",
            frame = FALSE,
            legend.outside = TRUE,
            bg.color = "grey70",
            legend.hist.width = 5,
            fontfamily = "Verdana")
```





